---
title: "Developing Data Products<br>Diamonds Price Prediction"
author: "Ioan Catana"
date: "July 30, 2018"
output: ioslides_presentation
---

```{r knitr_setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

## Introduction
This presentation is part of the Developing Data Products Coursera.org course project submission.

It is an R Presentation generated with RStudio.

The Shiny application pitched by this presentation is at:
[https://icatana.shinyapps.io/predict_diamonds/](https://icatana.shinyapps.io/predict_diamonds/)

The Shiny application source code is available at: [https://github.com/icatana/DDP-Course-Project](https://github.com/icatana/DDP-Course-Project)

## Goal

The goal of the application is to predict a diamond price using a prediction linear model. 

Also we will try to understand how different factors may influence the price of a diamond.

To address that we use the `diamonds` dataset containing the prices and other attributes of almost 54,000 diamonds.

To make interpretation easier we choose a random sample of maximum ten thousand entries from the original dataset.

## Plotting the Data

We plot the diamond weight(`carat`) vs. price in dollars (`price`) along the x/y axes. 
We also plot the prediction model in red.

```{r diamonds_plot, echo = FALSE}
library(plotly)
data(diamonds)
set.seed(112233)
small_diamonds <- diamonds[sample(nrow(diamonds), 1000), ]
fit <- lm(price ~ carat, data = small_diamonds)
plot_ly(small_diamonds, x = ~carat, y = ~price, color = I("blue"),
        text = ~paste("Carat: ", carat, '<br>Price:', price, ' $'),
        type = "scatter", mode = "markers", name = "Real Data") %>%
  add_lines(x = ~carat, y = fitted(fit), color = I("red"), name = "Prediction Model") %>%
  layout(title = "Diamonds Dataset")
```

## Code Sample 

The following is a sample code from Shiny application 'server.R':
```{r server_code, echo = TRUE, eval = FALSE}
  # Calculate price prediction based on linear model and input carat
  diamond_pred <- reactive({
    diamond_model <- lm(price ~ carat, data = small_diamonds())
    predict(diamond_model, newdata = data.frame(
      carat = input$vcarat))
  })
  # Draw the plot Diamond price vs. carat
  output$diamondsPlot <- renderPlot({
    ggplot(small_diamonds(), aes(x=carat,y=price)) +
      geom_point(color='blue',fill='blue') +
      geom_smooth(method = "lm", color = "red", se = FALSE) +
      geom_point(aes(x=input$vcarat, y=diamond_pred()), 
                 color='red', fill='red', size = 4)
    
  })
```
